---
output: github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# SwapPricer

<!-- badges: start -->
[![Project Status:Active](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)

<!-- badges: end -->

The goal of SwapPricer is to allow you to price a book of interest rate swaps (IRS).
IRS is the most traded over the counter financial derivative in the world and now
you can easily price it in R. More details can be found [here](https://en.wikipedia.org/wiki/Interest_rate_swap)

## Installation

<!-- You can install the released version of SwapPricer from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("SwapPricer") -->
<!-- ``` -->

You can for now just install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("DavideMagno/SwapPricer")
```
## Example

The main function of the toolbox is `SwapPortfolioPricing` which uses just three
inputs:

1) A table with the characteristics of the swap

2) The date at which the swaps are being priced

3) A table with the discounting factor curve

Examples of items 1 and 3 have been provided with the package.

```{r example}
library(SwapPricer)
SwapPortfolioPricing(SwapPricer::swap.basket, lubridate::ymd(20190414), SwapPricer::df.table)
```

This function returns a table that can be easily used for reporting like below

```{r cars, echo = FALSE}
SwapPricer::SwapPortfolioPricing(SwapPricer::swap.basket, lubridate::ymd(20190414), SwapPricer::df.table) %>% 
  dplyr::mutate_at(dplyr::vars("clean.mv","dirty.mv","accrual.pay","accrual.receive","pv01"),
                   formattable::comma) %>%
  dplyr::mutate_at(dplyr::vars("par"), formattable::percent) %>%
  knitr::kable(caption = "Pricing results", "html") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"))
```

## Learning More ##

The vignettes are a great place to learn more about SwapPricer.

If you are installing from github and want the vignettes, you'll need to run the following 
commands first:

```{r, eval = FALSE}
devtools::install_github("DavideMagno/SwapPricer", build_vignettes = TRUE)
library(SwapPricer)
```

You can then access them from R with the `vignette()` command.

