---
title: "Intro to SwapPricer"
author: "Davide Magno"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro_to_SwapPricer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `SwapPricer` package allows you to price a book of interest rate swaps (IRS).
IRS is the most traded over the counter financial derivative in the world and now
you can easily price it in R.  

Please note that as at version 0.2.0 the toolbox is able to price using a 
one-curve framework

**We are working to improve the toolbox in the next releases including OIS discounting**

Two main data structures are needed in order to run the SwapPriceR toolbox:

```{r, echo = FALSE}
`%>%` <- magrittr::`%>%`
`%<>%` <- magrittr::`%<>%`
```

**Swap Structure**

A swap can be defined as a list that includes the following needed fields:

* **currency** (character) as per the ISO 4217 standard (only `"EUR"` is currently active)
* **notional** (numeric)
* **start.date** (Date)
* **maturity.date** (Date)
* **strike** (numeric)
* **type** (string) either *payer* or *receiver*. It is not case-sensitive.
* **standard** (logical) 

Only if the field *standard* is set to FALSE then we need three additional fields

* **time.unit** (list) with two sub-fields: *pay* and *receive* and numerical data as per the number of months for the frequency of the leg (ie. monthly would have a time.unit of 1, quarterly of 3, semiannual of 6 and annual of 12)
* **dcc** (list) with two sub-fields: *pay* and *receive* and numerical data as per the RQuantLib helper
* **calendar** (character) the holiday's calendar as per the RQuantLib helper

The following one is an example of valid setup for both a standard swap and a non standard one.

```{r define_swap}
swap_standard <- list(currency = "EUR",
                      notional = 10000000,
                      start.date = lubridate::ymd(20070119),
                      maturity.date = lubridate::ymd(20320119),
                      strike = 0.00059820,
                      type = "receiver",
                      standard = TRUE)

swap_nonstandard <- list(currency = "EUR",
                         notional = 10000000,
                         start.date = lubridate::ymd(20070119),
                         maturity.date = lubridate::ymd(20320119),
                         strike = 0.00059820,
                         type = "receiver",
                         standard = FALSE,
                         time.unit = list(pay = 6, receive = 12),
                         dcc = list(pay = 1, receive = 6),
                         calendar = "TARGET")
```

Alternatively (and possibly more conveniently), you can parametrise your swap 
portfolio through Excel in a *csv* file like the following one:

```{r Get_Swaps_Data, warning=FALSE, message=FALSE}
SwapPricer::swap.basket %>% 
  dplyr::mutate_at(dplyr::vars("notional"), ~formattable::comma(., digits = 0)) %>%
  dplyr::mutate_at(dplyr::vars("strike"), formattable::percent) %>% 
  knitr::kable(caption = "Swap portfolio", "html", align = "c") %>% 
  kableExtra::scroll_box(width = "690px")
```

**Another important thing to note is that if you upload the csv file, the dates
must be in a dmy format. We are working on relaxing this constraint, but we 
haven't found the most suitable toolbox to help us (yes, we also tried `anytime`
and we were not fully impressed)**

This input will then be treated by a function in the toolbox to put in the list
format we have specified above.

**Discount Factor**

This has to be sourced from a third party provider (...for now...). 
In this case we downloaded the following data from Bloomberg.

```{r Get_DF_Data, warning=FALSE, message=FALSE, echo=FALSE}
SwapPricer:::ir.curve %>% 
  knitr::kable(caption = "Input from Bloomberg", "html") %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  kableExtra::scroll_box(height = "500px")
```

From this data we extracted the two most important columns, which will be used in the code of the toolbox:

* **t2m** which is the time to maturity of the different pillar points in the curve
* **df** the discount factor calculated for each pillar point

This data has to be stored in a tibble called **discount**. This tibble has to be
part of a list which will also contain the code of the curve currency.

Here's the structure of an example of interest rate list:

```{r, echo = FALSE}
str(SwapPricer::EUR.curves)
```

Now you have the two most important inputs that are needed to run your pricer. It is important to note that the interest rate curves' lists have to cover all the currencies present in the swap portfolio.


```{r}
today <- lubridate::ymd(20190415)

result <- SwapPricer::SwapPortfolioPricing(SwapPricer::swap.basket, today, SwapPricer::EUR.curves, SwapPricer::GBP.curves, SwapPricer::USD.curves) 
```

```{r, echo = FALSE}
result %>% 
  dplyr::mutate_at(dplyr::vars("clean.mv","dirty.mv","accrual.pay","accrual.receive","pv01"),
                   formattable::comma) %>%
  dplyr::mutate_at(dplyr::vars("par"), formattable::percent) %>%
  knitr::kable(caption = "Pricing results", "html") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"))
```

