---
title: "Introduction to SwapPricer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Define_data_structures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
The `SwapPricer` package allows you to price a book of interest rate swaps (IRS).
IRS is the most traded over the counter financial derivative in the world and now
you can easily price it in R.  

**Please note that as at version 0.1.0 the toolbox is not able to price:**

**1) Non EUR swaps**

**2) EUR swaps that do not pay quarterly on the variable rate**

** Another important thing to note is that if you upload the csv file, the dates
must be in a dmy format. We are working on relaxing this constraint, but we 
haven't found the most suitable toolbox to help us (yes, we also tried anytime
and we were not fully impressed) **

**We are working to solve all these issues in the next releases**

Two main data structures are needed in order to run the SwapPriceR toolbox:

```{r, echo = FALSE}
`%>%` <- magrittr::`%>%`
`%<>%` <- magrittr::`%<>%`
```

**Swap Structure**

A swap can be defined as a list that includes the following needed fields:

* **currency** (character) as per the ISO 4217 standard (only `"EUR"` is currently active)
* **notional** (numeric)
* **start.date** (Date)
* **maturity.date** (Date)
* **strike** (numeric)
* **type** (string) either *payer* or *receiver*. It is not case-sensitive.
* **standard** (logical) 

Only if the field *standard* is set to FALSE then we need three additional fields

* **time.unit** (list) with two sub-fields: *pay* and *receive* and numerical data as per the number of months for the frequency of the leg (ie. monthly would have a time.unit of 1, quarterly of 3, semiannual of 6 and annual of 12)
* **dcc** (list) with two sub-fields: *pay* and *receive* and numerical data as per the RQuantLib helper
* **calendar** (character) the holiday's calendar as per the RQuantLib helper

The following one is an example of valid setup for both a standard swap and a non standard one.

```{r define_swap}
swap_standard <- list(currency = "EUR",
                      notional = 10000000,
                      start.date = lubridate::ymd(20070119),
                      maturity.date = lubridate::ymd(20320119),
                      strike = 0.00059820,
                      type = "receiver",
                      standard = TRUE)

swap_nonstandard <- list(currency = "EUR",
                         notional = 10000000,
                         start.date = lubridate::ymd(20070119),
                         maturity.date = lubridate::ymd(20320119),
                         strike = 0.00059820,
                         type = "receiver",
                         standard = FALSE,
                         time.unit = list(pay = 6, receive = 12),
                         dcc = list(pay = 1, receive = 6),
                         calendar = "TARGET")
```



Alternatively (and possibly more conveniently), you can parametrise your swap 
portfolio through Excel in a *csv* file like the following one:

```{r Get_Swaps_Data, warning=FALSE, message=FALSE}
SwapPricer::swap.basket %>% 
  dplyr::mutate_at(dplyr::vars("notional"), ~formattable::comma(., digits = 0)) %>%
  dplyr::mutate_at(dplyr::vars("strike"), formattable::percent) %>% 
  knitr::kable(caption = "Swap portfolio", "html", align = "c") %>% 
  kableExtra::scroll_box(width = "690px")
```
This input will then be treated by a function in the toolbox to put in the list
format we have specified above.

**Discount Factor**

This has to be sourced from a third party provider (...for now...). 
In this case we downloaded the following data from Bloomberg.

```{r Get_DF_Data, warning=FALSE, message=FALSE, echo=FALSE}
SwapPricer:::ir.curve %>% 
  knitr::kable(caption = "Input from Bloomberg", "html") %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  kableExtra::scroll_box(height = "500px")
```

** Please note that for now we are not using dual curve construction and CSA
pricing. This is going to be part of a future release. We have decided to 
release the package anyway for you to familiarize with the package functions
first.**

From this data we extracted the two most important columns, which will be used in the code of the toolbox:

* **t2m** which is the time to maturity of the different pillar points in the curve
* **df** the discount factor calculated for each pillar point

The following code was used to calculate these two columns

```{r, warning = FALSE}
today <- lubridate::ymd(20190414)

SwapPricer::df.table %>% 
  knitr::kable(caption = "Discount Factor Curve", "html") %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  kableExtra::scroll_box(height = "500px")
```
Now you have the two most important inputs that are needed to run your pricer.

```{r}
result <- SwapPricer::SwapPortfolioPricing(SwapPricer::swap.basket, today, SwapPricer::df.table) 
```

```{r, echo = FALSE}
result %>% 
  dplyr::mutate_at(dplyr::vars("clean.mv","dirty.mv","accrual.pay","accrual.receive","pv01"),
                   formattable::comma) %>%
  dplyr::mutate_at(dplyr::vars("par"), formattable::percent) %>%
  knitr::kable(caption = "Pricing results", "html") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"))
```

